Reproducible Research Project 1
===============================

```{r, opts_chunk$set(echo = TRUE)}
```

#Loading and preprocessing data
```{r}
library(dplyr)
library(ggplot2)
download.file(url = "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip",
              destfile = "raw_data.zip")
unzip("raw_data.zip")
data <- read.csv("activity.csv", na.strings = "NA")
str(data)

#Convert date from a factor to a date
data$date <- as.Date(data$date, format = "%Y-%m-%d")
```

#What is the mean total number of steps taken per day
```{r}
#First, calculate the total number of steps per day and store in total dataframe
total <- data %>%
         group_by(date) %>%
         summarise(tot = sum(steps)) %>%
         na.omit()

hist(total$tot, xlab = "Steps", main = "Distribution of The Total Number of
     Steps Taken Each Day")

#Mean number of steps per day:
mean(total$tot)

#Median number of steps per day:
median(total$tot)
```

- The mean number of steps per day is `r mean(total$tot)`
- The median number of steps per is `r median(total$tot)`

#What is the average daily activity pattern
```{r}
interval <- data %>%
            group_by(interval) %>%
            na.omit(steps) %>%
            summarise(avg = mean(steps))

plot(interval$interval, interval$avg, type = "l",
        main = "Average Daily Activity Pattern",
        xlab = "Interval", ylab = "Average Number of Steps")

#5 minute interval with the maximum number of average steps
max(interval$avg)
```

#Imputing missing values
```{r}
#Number of rows with missing values:
sum(is.na(data))

#Fill in the missing data with the mean number of steps taken for each interval
fill <- merge(data, interval, by = "interval")
fill$steps <- ifelse(is.na(fill$steps), fill$avg, fill$steps)

total_na_filled <- fill %>%
                        group_by(date) %>%
                        summarise(total = sum(steps))

hist(total_na_filled$total, xlab = "Steps", main = "Distribution of The Total Number of
     Steps Taken Each Day with Missing Values Imputed")
        
#Averge of the total number of steps per day with imputed values:
mean(total_na_filled$total)

#Median of the total number of steps per day with imputed values:
median(total_na_filled$total)
```

Using the mean for each interval for imputation does not change the mean (still `r mean(total_na_filled$total)`)
when calculated without imputation (specifically, omitting the missing values). The median, however, is now equal to the mean.

#Weekdays vs. Weekends
```{r}
fill$date <- as.Date(fill$date, format = "%Y-%m-%d")
fill$weekday <- as.factor(ifelse(weekdays(fill$date) == "Saturday" | weekdays(fill$date) == 
                        "Sunday", "Weekend", "Weekday"))

ggplot(fill, aes(x = interval, y = avg)) +
        geom_line() +
        facet_grid(weekday ~ .) +
        ylab(label = "Average Number of Steps") +
        xlab(label = "Interval")
```
There is very little, if any, difference in the average number of steps taken on the weekends and weekdays, at least as the data has been calculated here. 